---
title: "Wine Quality Prediction"
author: "Tanmay Ambegaokar, Bhuimka Mallikarjun , Kundana Chowdary Cherukuri, Jagadeeshwar Kalyanapu"
date: "`r Sys.Date()`"
# date: ""
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
library(ezids)
library(tidyverse) 
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

**Loading the Dataset**
```{r load, include=TRUE, results='asis'}

wine=read.csv("winequalityN.csv")
xkabledplyhead(wine)
xkabledplytail(wine)
str(wine)
```



We have succesfully loaded the dataset. We can see that there are 6497 observations and 13 variables. Next lets look at the summary statistics. 

**SUMMARY STATISTICS**
```{r summary, include=TRUE, results='asis'}
xkablesummary(wine)
```

A quick look at the summary tells us the inter-quartile ranges and maximum and minimum values for each variable.

*Observations*

2) There are also some NA values which we need to remove.
3) There might be duplicates in the dataset.

**CLEANING THE DATA**

**Removing Duplicates**
```{r cleanduplicates, include=TRUE, results='asis'}
wine <- unique(wine)
str(wine)
```

Duplicates can cause issues later with biases. Thus, we have removed the duplicate values. Now, the dataset has 5329 observations.

**Removing NAs**
```{r removeNA, include=TRUE, results='asis'}
wine <- na.omit(wine)
str(wine)
xkablesummary(wine)
```


After, removing the NAs, we are finally left with 5295 observations. The dataset is now clean.

Now, the data is clean. We have NAs, and duplicates. We can now start our exploratory data analysis. 

*Observations*

1) There is a big difference between maximum value and Q3 values. This means that there are a quite a lot of outliers. Let's visualize this using plots.

